\documentclass[11pt]{article}


\def\showauthornotes{1}
\def\showdraftbox{0}

%%%%%%%%% Authornotes
\newcommand{\deeksha}{\Authornote{Deeksha}}
\newcommand{\todo}{\Authornote{TODO}}
\input{macros.tex}

\usepackage[
backend=biber,
backref=true,
backrefstyle=none,
date=year,
doi=false,
giveninits=true,
hyperref=true,
maxbibnames=10,
sortcites=false,
style=alphabetic,
url=false, 
]{biblatex}
\addbibresource{papers.bib}
\addbibresource{adaptive.bib}

\begin{document}
	
	
	\title{Decremental $(1+\epsilon)$-Approximate Maximum Eigenvector:\\Dynamic Power Method}
	\author{Deeksha Adil\thanks{Supported by Dr. Max R\"ossler, the Walter Haefner Foundation and the ETH Z\"urich Foundation}\\Institute for Theoretical Studies\\ETH Zürich\\deeksha.adil@eth-its.ethz.ch \and Thatchaphol Saranurak\thanks{Supported by NSF grant CCF-2238138.}\\Department of Computer Science\\University of Michigan\\thsa@umich.edu}
	\maketitle
	
	
	\maketitle
	
	
	\pagenumbering{gobble}
	\begin{abstract}
		We present a dynamic algorithm for maintaining $(1+\epsilon)$-approximate maximum eigenvector and eigenvalue of a positive semi-definite matrix $A$ undergoing \emph{decreasing} updates, i.e., updates which may only decrease eigenvalues. Given a vector $v$ updating $A\gets A-vv^{\top}$, our algorithm takes $\tilde{O}(\mathrm{nnz}(v))$ amortized update time, i.e., polylogarithmic per non-zeros in the update vector.
		
		Our technique is based on a novel analysis of the influential power method in the dynamic setting. The two previous sets of techniques have the following drawbacks (1) algebraic techniques can maintain exact solutions but their update time is at least polynomial per non-zeros, and (2) sketching techniques admit polylogarithmic update time but suffer from a crude additive approximation.
		
		Our algorithm exploits an oblivious adversary. 
        Interestingly, we show that any algorithm with polylogarithmic update time per non-zeros that works against an adaptive adversary and satisfies an additional natural property would imply a breakthrough for checking psd-ness of matrices in $\tilde{O}(n^{2})$ time, instead of $O(n^{\omega})$ time.
	\end{abstract}
	
	\newpage
	\pagenumbering{gobble}
	\tableofcontents
	\newpage
	
	\pagenumbering{arabic}
	\input{intro_new}
	\input{Preliminaries}
	\input{MaxEVDecNew}
	
	\input{conclusion}
	 \newpage
	\printbibliography
	
	
	 \newpage
	\appendix
	\input{dynSDP}
        \input{Appendix}
        
\end{document}
