\section{Connections to Dynamic Positive Semi-definite Programs}
\label{sec:dyn psd}
This section discusses connections between our \Cref{prob:dyn} and the dynamic versions of positive semi-definite programs. Using this connection, we conclude that \Cref{thm:upper} implies a dynamic algorithm for a special case of the dynamic covering SDP problem. 

We first define packing and covering semi-definite programs (SDPs).\footnote{Some papers \cite{jambulapati2021positive} flip our definition of packing and covering SDPs by considering their dual form.}


\begin{definition}[Packing/Covering SDP] Let $\CC, \AA_i$'s for $i = 1,2,...,m$ be $n\times n$ symmetric PSD matrices and $b_i$'s denote positive real numbers. The packing SDP problem asks to find
\[
\max_{\YY\succeq 0} \Tr[\CC\YY]  \quad \text{s.t.  } \Tr[\AA_i\YY]\leq b_i, \quad\forall i = 1,\cdots, m.
\]
The covering SDP problem asks to find
\[
\min_{\YY\succeq 0} \Tr[\CC\YY]  \quad \text{s.t.  } \Tr[\AA_i\YY]\geq b_i, \quad\forall i = 1,\cdots, m.
\]
\end{definition}
Note that when the matrices $\CC$ and $\AA_i$'s are all diagonal matrices, the packing and covering SDP problems above are precisely the well-studied packing and covering LP problems, respectively. Near-linear time algorithms for $(1+\eps)$-approximately solving packing and covering LPs are very well-studied in a long line of work, some of which are~\cite{allen2015nearly,allen2019nearly,wang2016unified,quanrud2020nearly}, and these problems have many applications, such as in graph embedding~\cite{plotkin1995fast}, approximation algorithms~\cite{luby1993parallel,trevisan1998parallel}, scheduling~\cite{plotkin1995fast}, to name a few.

\paragraph{Dynamic LPs.}
Near-optimal dynamic algorithms for packing and covering LPs were shown in~\cite{bhattacharya2023dynamic}. The paper studies two kinds of updates -- {\it restricting} and {\it relaxing} updates.
Restricting updates can only shrink the feasible region. In contrast, relaxing updates can only grow the feasible region.
In \cite{bhattacharya2023dynamic}, the authors gave a deterministic algorithm that can maintain a $(1+\epsilon)$-approximate solution to either packing and covering LPs that undergo only restricting updates or only relaxing updates in total time $\Otil(N/\epsilon^3 + t/\epsilon)$, where $N$ is the total number of nonzeros in the initial input and the updates, and $t$ is the number of updates. Hence, this is optimal up to logarithmic factors. 

A natural question is whether one can generalize the near-optimal dynamic LP algorithms with polylogarithmic overhead by \cite{bhattacharya2023dynamic} to work with SDPs since SDPs capture many further applications such as maximum cuts~\cite{iyengar2011approximating,klein1996efficient}, Sparse PCA~\cite{iyengar2011approximating}, sparsest cuts~\cite{iyengar2010feasible}, and  balanced separators~\cite{orecchia2012approximating}, among many others.

\paragraph{Static SDPs.}
Unfortunately, the algorithms for solving packing and covering SDPs are much more limited, even in the static setting. Near-linear time algorithms are known only for \emph{covering} SDPs when the cost matrix $\CC = \II$ is the identity \cite{peng2012faster,allen2016using}. 

The fundamental barrier to working with general psd matrix $\CC$ in covering SDPs is that it is as hard as approximating the minimum eigenvalue of $\CC$ (consider the program $\max_{\YY\succeq 0} \Tr[\CC\YY]$ such that $\Tr[\YY]\leq 1$). To the best of our knowledge, near-linear time algorithms for approximating the minimum eigenvalue assume a near-linear-time solver for $\CC$, i.e., to compute $\CC^{-1}\xx$ in the near-linear time given $\xx$. This can be done, for example, by applying the power method to $\CC^{-1}$. When $\CC^{-1}$ admits a fast solver, sometimes one can approximately solve a covering SDP fast, such as for spectral sparsification of graphs~\cite{lee2017sdp}, and the max-cut problem~\cite{arora2007combinatorial,trevisan2009max}.

For packing SDPs, there is simply no near-linear time algorithm known. An algorithm for approximately solving packing SDPs and even the generalization to mixed packing-covering SDPs was claimed in \cite{jambulapati2021positive}, but there is an issue in the convergence analysis even for pure packing SDPs. Fast algorithms for this problem, hence, remain open.

\paragraph{Dynamic SDPs.}
Since near-linear time static algorithms are prerequisites for dynamic algorithms with polylogarithmic overhead, we can only hope for a dynamic algorithm for covering SDPs when $\CC$ is an identity. 
Below, we will show that our algorithm for \Cref{thm:upper} implies a dynamic algorithm for maintaining the covering SDP solution when there is a single constraint and the updates are restricting. This follows because this problem is equivalence to \Cref{prob:dyn}.



We first define the dynamic covering problem with a single constraint under restricting updates.

\begin{problem}[Covering SDP with a Single Matrix Constraint under Restricting Updates]\label{prob:dynsdp} Given $\AA_0\succeq 0$, an accuracy parameter $\epsilon>0$, and an online sequence of vectors $\vv_1,\vv_2,\cdots,\vv_T$ that update $\AA_t \gets \AA_t -\vv_t\vv_t^{\top}$ such that $\AA_t\succeq 0$. The problem asks to explicitly maintain, for all $t$, an $(1+\eps)$-approximate optimal value $\nu_t$ of the SDP, i.e.,  $$\nu_t \leq (1+\epsilon)OPT_t \defeq {\min}_{\YY\succeq 0, \Tr[\AA_t\YY]\geq 1}  \Tr[\YY].$$
Furthermore, given a query request, return a matrix $\QQ^{(t)}$ where $\YY^{(t)} = \QQ^{(t)}{\QQ^{(t)}}^{\top} \in \mathbb{R}^{n\times n}$ is a $(1+\eps)$-approximate optimal solution, i.e., 
    $$\Tr[\AA_t\YY^{(t)}]\geq 1\text{ and }\Tr[\YY^{(t)}]\leq \left(1+\epsilon\right) OPT_t.$$
\end{problem}
\Cref{prob:dynsdp} is equivalent to \Cref{prob:dyn} in the following sense: given an algorithm for \Cref{prob:dyn}, we can obtain an algorithm for \Cref{prob:dynsdp} with the same total update time and optimal query time. Conversely, given an algorithm for \Cref{prob:dynsdp}, we can obtain an algorithm for \Cref{prob:dyn} in the \emph{eigenvalue-only} version with the same total update time.

\begin{proposition}\label{prop:equiv sdp}
The following holds,
\begin{enumerate}
    \item Given an algorithm for \Cref{prob:dyn} with update time $\mathcal{T}$, there is an algorithm for Problem~\ref{prob:dynsdp} with update time 
 $\mathcal{T}$ and query time $O(n)$, i.e., the time required to query $\QQ^{(t)}$ at any time $t$. 
    \item Given an algorithm for Problem~\ref{prob:dynsdp} with update time $\widetilde{\mathcal{T}}$, there is an algorithm for \Cref{prob:dyn} that only maintains the approximate eigenvalues with update time at most $\widetilde{\mathcal{T}}$.
\end{enumerate}
\end{proposition}
\begin{proof}
  Let us first characterize the solution to Problem~\ref{prob:dynsdp} at any $t$ for $\epsilon=0$. Since any feasible solution $\YY$ is PSD it can be characterized as, $\YY = \sum_{i=1}^n p_i \yy_i\yy_i^{\top}$ for some unit vectors $\yy_i$'s. Now $\Tr[\YY] = \sum_i p_i$ and
  \[
 1\leq  \Tr[\AA_t\YY] = \sum_{i=1}^n p_i \Tr[\AA_t \yy_i\yy_i^{\top}] = \sum_{i=1}^n p_i \yy_i^{\top}\AA_t\yy_i \leq \lambda_{\max}(\AA_t) \sum_{i=1}^n p_i.
  \]
  The above implies that for any feasible solution $\YY^{(t)}$, it must hold that $\sum_{i=1}^n p_i\geq 1/\lambda_{\max}(\AA_t)$, and equality holds iff $\yy_i$'s are the maximum eigenvector of $\AA_t$ for all $t$. Since the problem is asking to minimize $\sum_{i=1}^n p_i$, the solution must be $\YY^{(t)} = \frac{1}{\lambda_{\max}(\AA_t)}\uu\uu^{\top}$ where $\uu$ is the maximum eigenvector of $\AA_t$. Note that here $\QQ^{(t)} = \frac{1}{\sqrt{\lambda_{\max}(\AA_t)}}\uu$. 

  We now show the first part, by proving that at any $t$, given $\epsilon$, the solution to \Cref{prob:dyn} gives a solution to Problem~\ref{prob:dynsdp}. Let $\lambda_t$ and $\ww_t$ denote an $\epsilon/2$-approximate solution to \Cref{prob:dyn} for some $t$. Consider the solution $\QQ^{(t)} = \frac{1}{\sqrt{(1-\epsilon/2)\lambda_t}}\ww_t$ which gives $\YY^{(t)} = \frac{1}{(1-\epsilon/2)\lambda_t}\ww_t\ww_t^{\top}$. Then,
  \[\Tr[\AA_t\YY^{(t)}] = \frac{1}{(1-\epsilon/2)\lambda_t} \ww_t^{\top}\AA_t\ww_t \geq (1-\epsilon/2)\frac{\lambda_{\max}(\AA_t)}{(1-\epsilon/2)\lambda_{\max}(\AA)}\geq 1.\]
Therefore, $\YY^{(t)}$ satisfies the constraints of Problem~\ref{prob:dynsdp}. Next we look at the objective value, $\nu_t = \Tr[\YY^{(t)}] = \frac{1}{(1-\epsilon/2)\lambda_t} \leq \frac{1}{(1-\epsilon/2)^2\lambda_{\max}(\AA_t)} \leq (1+\epsilon) OPT_t.$ We can maintain $\QQ^{(t)}$ by just maintaining $\ww_t,\lambda_t$ which requires no extra time. The time required to obtain $\QQ^{(t)}$, which is the query time, from $\ww_t$ and $\lambda_t$ is at most $O(nnz(\ww_t))=O(n)$ and the value of $\nu_t$ can be obtained in $O(1)$ time. 

To see the other direction, consider the solution $\QQ^{(t)}$, $\nu_t$ of Problem~\ref{prob:dynsdp}. We can set $\lambda_t = \frac{1}{\nu_t}$. This implies that,
\[
\lambda_t = \frac{1}{\nu_t}\geq \frac{\lambda_{\max}(\AA_t)}{(1+\epsilon)} \geq (1-\epsilon)\lambda_{\max}(\AA_t),
\]
as required. In this case, we do not require any extra time.
\end{proof}

By plugging \Cref{thm:upper} into \Cref{prop:equiv sdp}, we conclude the following.
\begin{corollary} There is a randomized algorithm for \Cref{prob:dynsdp} under a sequence of $T$ restricting updates, that given $n, \AA_0$ and $\epsilon>1/n$ as input, with probability at least $1-1/n$ works against an oblivious adversary in total update time
\[
 O\left(\frac{\log^{3}n \log^6 \frac{n}{\epsilon}\log \frac{\lambda_{\max}(\AA_0)}{\lambda_{\max}(\AA_T)}}{\epsilon^{4}}\left(nnz(\AA_0) + \sum_{i=1}^T nnz(\vv_i)\right) \right),
 \]
and query time $O(n)$.
\end{corollary}