In order to relax the notion of satisfiability, we first consider CSPs on $U_d$
for some $d$ and represent CSPs via polynomials. Let $\Gm$ be a constraint 
language over $U_d$. Every constraint $\ang{\bs,\rel}$ of an instance 
$\cP=(V,U_d,\cC)$ of $\CSP(\Gm)$ is represented by a polynomial 
$P_\rel(\bs)$ that represents the characteristic function $f_\rel$ of
$\rel$:
%
\[
f_\rel(\bs)=\left\{\begin{array}{ll}
  \ld_0, & \text{if $\rel(\bs)$ is true,}\\
  \ld_1, & \text{otherwise}.
\end{array}\right.
\]
%
We note that our choice of the polynomial representation is somewhat arbitrary
but other choices lead to the same results. (For instance, \cite{AKS19:jcss}
studied the $d=2$ case and used $\ld_1$ to represent true.) 

An operator $A$ on a Hilbert space $\cH$ is a \emph{normal operator of order
$d$} if $A$ is normal and $A^d=I$. Operator assignment to the instance $\cP$ on a Hilbert space $\cH$ is a mapping 
that assigns to every variable from $V$ an operator $A_v$ on $\cH$ such that
\begin{itemize}
\item[(a)]
$A_v$ is a normal operator of order $d$ for every $v\in V$;
\item[(b)]
the operators $A_{v_1}\zd A_{v_r}$ pairwise commute
for every constraint  $\ang{(\vc vr),\rel}\in\cC$.
\end{itemize}

\noindent
We call an operator assignment $\{A_v\}$ \emph{satisfying for $\cP$} if 
$P_\rel(A_{v_1}\zd A_{v_r})=I$ for every constraint $\ang{(\vc vr),\rel}\in\cC$.
%
Let $\cP$ be a CSP instance. 
Following the terminology of Atserias et al.~\cite{AKS19:jcss}, we say that
$\cP$ has a \emph{satisfiability gap of the first kind} if $\cP$ is not
satisfiable over $U_d$ but is satisfiable by an operator assignment on 
a finite-dimensional Hilbert space. Similarly, we say that 
$\cP$ has a \emph{satisfiability gap of the second kind} if $\cP$ is not
satisfiable over $U_d$ but is satisfiable by an operator assignemnt on an
infinite-dimensional Hilbert space. Finally, we say that 
$\cP$ has a \emph{satisfiability gap of the third kind} if $\cP$
is not satisfiable on finite-dimensional Hilbert spaces but is satisfiable by an
operator assignment on 
an infinite-dimensional Hilbert space.
%
We say that $\CSP(\Gm)$ has a satisfiablity gap of the $i$-th
kind, $i=1,2,3$, if there is at least one instance $\cP\in\CSP(\Gm)$ witnessing
such a gap.
%
By definition, a gap of the first kind implies a gap of the second kind. Also, a
gap of the third kind implies a gap of the second kind. Put differently, if
$\Gm$ has no gap of the second kind then it has no gap of the first or third
kind either.

We shall repeatedly use the following simple lemma, proved in Appendix~\ref{app:lemma}.

\begin{lemma}\label{lem:lemma-3}
Let $\vc xr$ be variables, let $\vc Qm,Q$ be polynomials in $\zC[\vc xr]$,
and let $\cH$ be a Hilbert space. If every assignment over $U_d$ that 
satisfies the equations $Q_1=\dots=Q_m=0$ also satisfies the equation
$Q=0$, then every fully commuting operator assignment on $\cH$ that 
satisfies the equations $Q_1=\dots=Q_m=0$ also satisfies the equation
$Q=0$.
\end{lemma}

\input{i-lemma}

